Automated Residual Plot Assessment with the R Package autovi and the Shiny App autovi.web

### 2024-11-26 autovi_paper.qmd
Join this sentence to the prev paragraph

However, as discussed in @li2024automated, the lineup protocol has significant limitations in large-scale applications due to dependence on human labor. Thus, a computer vision model was developed with an associated statistical testing procedure to automate the assessment of residual plots. This model takes a residual plot and a vector of auxiliary variables (such as the number of observations) as inputs and outputs the predicted visual signal strength (VSS). This strength estimates the distance between the residual distribution of the fitted regression model and the reference distribution assumed under correct model specification.

<!--Software for regression analysis tools is widely available. The Comprehensive R Archive Network (CRAN) [@hornik2012comprehensive] hosts a vast array of packages, many of which provide tools to diagnose models using residual plots. These packages can be broadly categorized into three groups: general-purpose, enhanced visual diagnostics, and visual diagnostics with statistical testing.-->

<!--General-purpose regression analysis tools are the most widely used group of packages. While not specifically designed for graphical diagnostics of residuals in linear regression, they include this functionality as part of a broader suite of statistical tools. For example, the `stats` package [@stats] offers standard diagnostic plots such as residuals vs. fitted values, quantile-quantile (Q-Q) plots, and residuals vs. leverage plots. Additional packages like `jtools` [@jtools], `olsrr` [@olsrr], `rockchalk` [@rockchalk], and `ggResidpanel` [@ggresidpanel] provide similar graphical diagnostics, often with alternative aesthetics or interactive features. All of these tools deliver the types of diagnostic plots outlined in the classical text by @cook1982residuals. However, as discussed in @li2024plot, relying solely on subjective assessments of these plots can lead to issues, such as over-interpreting random patterns as model violations.-->

<!-- The second group consists of enhanced visual diagnostics tools, which provide advanced aids for interpreting diagnostic plots. For instance, `ecostats` [@warton_global_2023] incorporates simulation envelopes into residual plots, while `DHARMa` [@dharma] compares empirical quantiles (0.25, 0.5, and 0.75) of scaled residuals to their theoretical counterparts. `DHARMa` is particularly focused on detecting model violations such as heteroscedasticity, incorrect functional forms, and issues specific to generalized linear and mixed-effect models, like over/under-dispersion. It also includes annotations from conventional tests, displayed as labels or text within the plot, to help avoid misinterpretation.-->

<!-- The third group, visual diagnostics with statistical testing, offers tools for formally testing patterns observed in diagnostic plots [@buja2009statistical]. A common method for visual testing of residual plots is the lineup protocol, where the true residual plot is embedded within a set of null plots generated by simulating residuals under the null hypothesis $H_0$ that the regression model is correctly specified. Observers are asked to identify the plot that appears most different from the others. If a significant proportion of observers correctly identify the true residual plot, it provides evidence against $H_0$, as the true residual plot should be indistinguishable from the null plots if all residuals are generated by the same process, according to @buja2009statistical. Packages in this group, such as `nullabor` [@nullabor], `HLMdiag` [@loy2014hlmdiag], and `regressinator` [@regressinator], enable users to compare observed residual plots with samples from null distributions, helping to quantify the significance of any detected patterns. -->

<!-- However, as discussed in @li2024automated, the lineup protocol has significant limitations in large-scale applications due dependence on human labor. Thus a computer vision model was developed with an associated statistical testing procedure to automate the assessment of residual plots. This model takes a residual plot and a vector of auxiliary variables (such as the number of observations) as inputs and outputs the predicted visual signal strength (VSS). This strength estimates the distance between the residual distribution of the fitted regression model and the reference distribution assumed under correct model specification.-->

To make the statistical testing procedure and trained computer vision model widely accessible, we developed the R package `autovi`, and a web interface, `autovi.web` to make it easy for users to automatically read their residual plots with the trained computer vision model.

>>>>

However, as discussed in @li2024automated, the lineup protocol has significant limitations in large-scale applications due to dependence on human labor. Thus, a computer vision model was developed with an associated statistical testing procedure to automate the assessment of residual plots. This model takes a residual plot and a vector of auxiliary variables (such as the number of observations) as inputs and outputs the predicted visual signal strength (VSS). This strength estimates the distance between the residual distribution of the fitted regression model and the reference distribution assumed under correct model specification.
<!---->
<!--Software for regression analysis tools is widely available. The Comprehensive R Archive Network (CRAN) [@hornik2012comprehensive] hosts a vast array of packages, many of which provide tools to diagnose models using residual plots. These packages can be broadly categorized into three groups: general-purpose, enhanced visual diagnostics, and visual diagnostics with statistical testing.-->
<!---->
<!--General-purpose regression analysis tools are the most widely used group of packages. While not specifically designed for graphical diagnostics of residuals in linear regression, they include this functionality as part of a broader suite of statistical tools. For example, the `stats` package [@stats] offers standard diagnostic plots such as residuals vs. fitted values, quantile-quantile (Q-Q) plots, and residuals vs. leverage plots. Additional packages like `jtools` [@jtools], `olsrr` [@olsrr], `rockchalk` [@rockchalk], and `ggResidpanel` [@ggresidpanel] provide similar graphical diagnostics, often with alternative aesthetics or interactive features. All of these tools deliver the types of diagnostic plots outlined in the classical text by @cook1982residuals. However, as discussed in @li2024plot, relying solely on subjective assessments of these plots can lead to issues, such as over-interpreting random patterns as model violations.-->
<!---->
<!-- The second group consists of enhanced visual diagnostics tools, which provide advanced aids for interpreting diagnostic plots. For instance, `ecostats` [@warton_global_2023] incorporates simulation envelopes into residual plots, while `DHARMa` [@dharma] compares empirical quantiles (0.25, 0.5, and 0.75) of scaled residuals to their theoretical counterparts. `DHARMa` is particularly focused on detecting model violations such as heteroscedasticity, incorrect functional forms, and issues specific to generalized linear and mixed-effect models, like over/under-dispersion. It also includes annotations from conventional tests, displayed as labels or text within the plot, to help avoid misinterpretation.-->
<!---->
<!-- The third group, visual diagnostics with statistical testing, offers tools for formally testing patterns observed in diagnostic plots [@buja2009statistical]. A common method for visual testing of residual plots is the lineup protocol, where the true residual plot is embedded within a set of null plots generated by simulating residuals under the null hypothesis $H_0$ that the regression model is correctly specified. Observers are asked to identify the plot that appears most different from the others. If a significant proportion of observers correctly identify the true residual plot, it provides evidence against $H_0$, as the true residual plot should be indistinguishable from the null plots if all residuals are generated by the same process, according to @buja2009statistical. Packages in this group, such as `nullabor` [@nullabor], `HLMdiag` [@loy2014hlmdiag], and `regressinator` [@regressinator], enable users to compare observed residual plots with samples from null distributions, helping to quantify the significance of any detected patterns. -->

<!-- However, as discussed in @li2024automated, the lineup protocol has significant limitations in large-scale applications due dependence on human labor. Thus a computer vision model was developed with an associated statistical testing procedure to automate the assessment of residual plots. This model takes a residual plot and a vector of auxiliary variables (such as the number of observations) as inputs and outputs the predicted visual signal strength (VSS). This strength estimates the distance between the residual distribution of the fitted regression model and the reference distribution assumed under correct model specification.-->
To make the statistical testing procedure and trained computer vision model widely accessible, we developed the R package `autovi`, and a web interface, `autovi.web` to make it easy for users to automatically read their residual plots with the trained computer vision model.
